---
title: "Basic coding and Regular expressions"
author: "Anoob Prakash"
date: "19 September 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Coding in R
Suppose x = 1.1, a = 2.2, and b = 3.3. Assign each expression to the value of the variable zand print the value stored in z.

a. $$x^{a^b}$$

```{r, echo=TRUE}
x <- 1.1
a <- 2.2
b <- 3.3

z <- x^a^b
print(z)
```

b. $$(x^{a})^b$$ 

```{r, echo=TRUE}
z <- (x^a)^b
print(z)
```

c. $$3x^3 + 2x^2 + 1$$

```{r, echo=TRUE}
z <- 3*x^3 + 2*x^2 + 1
print(z)
```

***
Using the rep and seq functions, create the following vectors:

a. (1,2,3,4,5,6,7,8,7,6,5,4,3,2,1)

```{r,echo=TRUE}
seq1 <- seq(1:8)
seq2 <- seq(from = 7, to =1)
z <- c(seq1, seq2)
print(z)

```
b. (1,2,2,3,3,3,4,4,4,4,5,5,5,5,5)

```{r,echo=TRUE}
seq3 <- seq(1:5)
vec1 <- c(1,2,3,4,5)
z <- rep(x = seq3, times = vec1)
print(z)
```
c. (5,4,4,3,3,3,2,2,2,2,1,1,1,1,1)
```{r,echo=TRUE}
seq4 <- seq(from = 5, to = 1)
vec1 <- c(1,2,3,4,5)
z <- rep(x = seq4, times = vec1)
print(z)
```
***

Create a vector of two random uniform numbers. In a spatial map, these can be interpreted as x and y coordinates that give the location of an individual (such as a marked forest tree in a plot that has been mapped). Using one of R’s inverse trigonometry functions (asin(), acos(), or atan()), convert these numbers into polar coordinates
```{r, echo=TRUE}
# background information for this conversion.
# To convert from Cartesian Coordinates (x,y) to Polar Coordinates (r,θ) :
# r = sqrt(x^2 + y^2) 
# θ = atan(y/x)


# Generating random unifrom number
xy <- runif(2)

# let r and theta be the two polar coordinates
r <- sqrt(xy[1]^2 + xy[2]^2)

theta <- atan(xy[2]/xy[1])

print(r)
print(theta)
```
***
Suppose that queue <- c("sheep", "fox", "owl", "ant") and that queue represents the animals that are lined up to enter Noah’s Ark, with the sheep at the front of the line. Using R expressions, update the queue successively as
a. the serpent arrives;
b. the sheep enters the ark;
c. the donkey arrives and talks his way to the front of the line;
d. the serpent gets impatient and leaves;
e. the owl gets bored and leaves;
f. the aphid arrives and the ant invites him to cut in line.
Finally, determine the position of the aphid in the line.

```{r, echo=TRUE}
# Creating the initial queue with the concatenate 
queue <- c("sheep", "fox", "owl", "ant")
queue <- c(queue, "serpent")
print(queue) # the serpent arrives in the queue
queue <- c(queue[-1])
print(queue) # the sheep enters the ark and leaves the queue
queue <- c("donkey",queue)
print(queue) # the donkey arrives and talks his way to the front of the line
queue <- c(queue[-5])
print(queue) # serpent leaves the queue
queue <- c(queue[-3])
print(queue) # bored owl leaves the queue
queue <- c(queue[queue!="ant"], "aphid","ant")
print(queue) # aphid cuts in the line with ant
aphidsPosition <- which(queue == "aphid") #using the which() to find the exact position of the aphid in the queue
print(aphidsPosition) # position of the aphid
```
***
Use R to create a vector of all of the integers from 1 to 100 that are not divisible by 2, 3, or 7.
```{r,echo=TRUE}

vec1 <- seq(from=1,to=100) # create vector from 1 to 100

vec2 <- vec1[(vec1%%2)!=0] # remove numbers not divisible by 2 

vec3 <- vec2[(vec2%%3)!=0] # remove numbers not divisible by 3 

vec <- vec3[(vec3%%7)!=0] # remove numbers not divisible by 7

print(vec)

# another method using the & sign
vec1 <- seq(from=1,to=100)
vectr <- vec1[(vec1%%2)!=0 & (vec1%%3)!=0 & (vec1%%7)!=0]
print(vectr)
```
***
***

# Regular Expression Puzzles
For these problems, use your plain text editor (BBedit, NotePadPro, or something else) to type in the problem text and use the search function to write a regular expression that gives the desired result. In your homework solution, provide the regular expression that works (there are several ways to solve each problem) within some plain text fencing on your markdown page, and add a bit of markdown text to explain what each element of your regular expression is doing. If you get stuck, give the solution that gets you as close as you can.
***
Q. The primary reason for using Excel to set up data frames is that people like to have the columns aligned. However, if there are not too many columns, it may be faster to do the job in a plain text editor first and align the columns with tabs. In your text editor, type in (or copy and paste from here) the following lines of text:
```
First String    Second      1.22      3.4
Second          More Text   1.555555  2.2220
Third           x           3         124
```
Don’t worry about how many tab spaces are needed to set this up, just make sure the columns are aligned. Now, using a single regular expression, transform these lines into what we need for a .csv file:
```
First String,Second,1.2,3.4
Second,More Text,1.55555,2.2220
Third,x,3,124
```
Ans:
```
Find:\t{1,}

Replace with:,
```
***
Q. A True Regex Story. I am preparing a collaborative NSF grant with a colleague at another university. One of the pieces of an NSF grant is a listing of potential conflicts of interest. NSF wants to know the first and last name of the collaborator and their institution.
Here are a few lines of my conflict list:
```
Ballif, Bryan, University of Vermont
Ellison, Aaron, Harvard Forest
Record, Sydne, Bryn Mawr
```
However, my collaborator asked me to please provide to her the list in this format:
```
Bryan Ballif (University of Vermont)
Aaron Ellison (Harvard Forest)
Sydne Record (Bryn Mawr)
```
Write a single regular expression that will make the change.
Ans:
```
Find:(\w*),\s(\w*),\s(.*)

Replace with:\2 \1 \(\3\)  #incase this does not complete the task, use the following:
Replace with (alt): \2 \1 [\3]
```
***
Q. A Second True Regex Story. A few weeks ago, at Radio Bean’s Sunday afternoon old-time music session, one of the mandolin players gave me a DVD with over 1000 historic recordings of old-time fiddle tunes.
The list of tunes (shown here as a single line of text) looks like this:
```
0001 Georgia Horseshoe.mp3 0002 Billy In The Lowground.mp3 0003 Cherokee Shuffle.mp3 0004 Walking Cane.mp3
```
Unfortunately, in this form, you can’t re-order the file names to put them in alphabetical order. I thought I could just strip out the leading numbers, but this will cause a conflict, because, for wildly popular tunes such as “Shove That Pig’s Foot A Little Further In The Fire”, there are multiple copies somewhere in the list.

All of these files are on a single line, so first write a regular expression to place each file name on its own line:
```
0001 Georgia Horseshoe.mp3
0002 Billy In The Lowground.mp3
0003 Cherokee Shuffle.mp3
0004 Walking Cane.mp3
```
Ans:
```
Find:.mp3\s

Replace with:.mp3\n
```
Now write a regular expression to grab the four digit number and put it at the end of the title:
```
Georgia Horseshoe_0001.mp3
Billy In The Lowground_0002.mp3
Cherokee Shuffle_0003.mp3
Walking Cane_0004.mp3
```
Ans:
```
Find:(\d{4})\s(.*)(\.mp3)

Replace with:\2_\1\3
```
Q. Here is a data frame with genus, species, and two numeric variables.
```
Camponotus,pennsylvanicus,10.2,44
Camponotus,herculeanus,10.5,3
Myrmica,punctiventris,12.2,4
Lasius,neoniger,3.3,55
```
Write a single regular expression to rearrange the data set like this:
```
C_pennsylvanicus,44
C_herculeanus,3
M_punctiventris,4
L_neoniger,55
```
Ans:
```
Find:(\w{1})\w+,(\w*),(.*),(d*)

Replace with:\1_\2,\4
```
Beginning with the original data set, rearrange it to abbreviate the species name like this:
```
C_penn,44
C_herc,3
M_punc,4
L_neon,55
```
Ans:
```
Find:(\w{1})\w+,(\w{4})\w+,(.*),(d*)

Replace with:\1_\2,\4
```