---
title: "Data Distributions"
author: "Anoob Prakash"
date: "3 October 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
***

####Long and wide data formats
Try converting the iris data set into the long format, with a column called “trait” to indicate sepal and petal length and width.

Once you have converted to the long format, calculate the average for each combination of species and trait.

```{r}
library(tidyverse)
longDat <- iris %>%
  gather(Sepal.Length:Petal.Width, key = "trait", value = "value") %>%
  arrange(Species)

head(longDat)

SpsTrait <- longDat %>%
  group_by(Species,trait) %>%
  summarize(meanValues = mean(value)) %>%
  arrange(Species,trait)
print(SpsTrait)
```
***
####Simulating and Fitting Data Distributions
This exercise teaches you how to compare a histogram of continuous (or integer) data to the probability density functions for different statistical distributions.

Set up a new .Rmd file for this exercise. Copy and paste the code below into different code chunks, and then read the text and run the code chunks one at a time to see what they do. You probably won’t understand everything in the code, but this is a good start for seeing some realistic uses of ggplot. We will cover most of these details in the next few weeks.

Once the code is in and runs, try reading in your own .csv file into a data frame with this code chunk:

```{r}
z <- read.table("MyDataFile.csv",header=TRUE,sep=",", stringsAsFactors=FALSE)
str(z)
summary(z)
print(z)
```
```
z <- rnorm(n=3000,mean=0.2)
z <- data.frame(1:3000,z)
names(z) <- list("ID","myVar")
z <- z[z$myVar>0,]
str(z)
summary(z$myVar)
head(z)