---
title: "Data Distributions"
author: "Anoob Prakash"
date: "3 October 2018"
output: html_document
---


####Long and wide data formats
Try converting the iris data set into the long format, with a column called “trait” to indicate sepal and petal length and width.

Once you have converted to the long format, calculate the average for each combination of species and trait.

```{r}
library(tidyverse)
longDat <- iris %>%
  gather(Sepal.Length:Petal.Width, key = "trait", value = "value") %>%
  arrange(Species)

head(longDat)

SpsTrait <- longDat %>%
  group_by(Species,trait) %>%
  summarize(meanValues = mean(value)) %>%
  arrange(Species,trait)
print(SpsTrait)
```
***
####Simulating and Fitting Data Distributions
The data used for fitting the distribution is from a study conducted by me on `moist deciduous trees` under particulate pollution of Western Ghats of Kerala, India. Out of the several variable studied, I am using the `relative water content` ([RWC](MyDataFile.csv)) as the variable for testing the effects of particulate pollution. :

```{r}
z <- read.table("MyDataFile.csv",header=TRUE,sep=",", stringsAsFactors=FALSE) # myVar is the RWC for the species studied
str(z)
summary(z)
print(z)
```

```{r}
library(ggplot2) # for graphics
library(MASS) # for maximum likelihood estimation
```

***
####Plot histogram of data
Plot a histogram of the data using ggplot. Rescaling the y axis of the histogram from counts to density, so that the area under the histogram equals 1.0.
```{r}
p1 <- ggplot(data=z, aes(x=myVar, y=..density..)) +
  geom_histogram(color="grey60",fill="cornsilk",size=0.2) 
print(p1)
```

***

####Add empirical density curve
Modifying the code to add in a kernel density plot of the data. This is an empirical curve that is fitted to the data. It does not assume any particular probability distribution, but it smooths out the shape of the histogram:
```{r}
p1 <-  p1 +  geom_density(linetype="dotted",size=0.75)
print(p1)

```

***

####Get maximum likelihood parameters for normal
Fitting a normal distribution to the data and grabbing the maximum likelihood estimators of the two parameters of the normal, the mean and the variance:

```{r}
normPars <- fitdistr(z$myVar,"normal")
print(normPars)
str(normPars)
normPars$estimate["mean"] # note structure of getting a named attribute
```

***

####Plot normal probability density
Let’s call the dnorm function inside ggplot’s stat_function to generate the probability density for the normal distribution. 

Note: Read about  stat_function in the help system to see how you can use this to add a smooth function to any ggplot. Note that we first get the maximum likelihood parameters for a normal distribution fitted to thse data by calling fitdistr. Then we pass those parameters (meanML and sdML to  stat_function:
```{r}
meanML <- normPars$estimate["mean"]
sdML <- normPars$estimate["sd"]

xval <- seq(0,max(z$myVar),len=length(z$myVar))

 stat <- stat_function(aes(x = xval, y = ..y..), fun = dnorm, colour="red", n = length(z$myVar), args = list(mean = meanML, sd = sdML))
 p1 + stat
```


***

####Plot exponential probability density
Now let’s use the same template and add in the curve for the exponential:
```{r}
expoPars <- fitdistr(z$myVar,"exponential")
rateML <- expoPars$estimate["rate"]

stat2 <- stat_function(aes(x = xval, y = ..y..), fun = dexp, colour="blue", n = length(z$myVar), args = list(rate=rateML))
 p1 + stat + stat2
```

***

####Plot uniform probability density
For the uniform, we don’t need to use fitdistr because the maximum likelihood estimators of the two parameters are just the minimum and the maximum of the data:
```{r}
stat3 <- stat_function(aes(x = xval, y = ..y..), fun = dunif, colour="darkgreen", n = length(z$myVar), args = list(min=min(z$myVar), max=max(z$myVar)))
 p1 + stat + stat2 + stat3
```

***

####Plot gamma probability density
```{r}
gammaPars <- fitdistr(z$myVar,"gamma")
shapeML <- gammaPars$estimate["shape"]
rateML <- gammaPars$estimate["rate"]

stat4 <- stat_function(aes(x = xval, y = ..y..), fun = dgamma, colour="brown", n = length(z$myVar), args = list(shape=shapeML, rate=rateML))
 p1 + stat + stat2 + stat3 + stat4
```

***

####Plot beta probability density
This one has to be shown in its own plot because the raw data must be rescaled so they are between 0 and 1, and then they can be compared to the beta.

```{r}
pSpecial <- ggplot(data=z, aes(x=myVar/(max(myVar + 0.1)), y=..density..)) +
  geom_histogram(color="grey60",fill="cornsilk",size=0.2) + 
  xlim(c(0,1)) +
  geom_density(size=0.75,linetype="dotted")

betaPars <- fitdistr(x=z$myVar/max(z$myVar + 0.1),start=list(shape1=1,shape2=2),"beta")
shape1ML <- betaPars$estimate["shape1"]
shape2ML <- betaPars$estimate["shape2"]

statSpecial <- stat_function(aes(x = xval, y = ..y..), fun = dbeta, colour="orchid", n = length(z$myVar), args = list(shape1=shape1ML,shape2=shape2ML))
pSpecial + statSpecial
```


From the above fittings, I found the `normal distribution to be a better fit` for the data on RWC for moist deciduous species under particulate pollution stress.